Trends of SDR in Thessaloniki

### 2023-06-09 Article_B140.Rmd
Commit 2023-06-09 13:17

---
title:         "Trends of SDR in Thessaloniki"

# 
# header-includes:
#   - \usepackage{caption}
#   - \usepackage{placeins}
#   - \captionsetup{font=small}
#   - \usepackage{setspace}\doublespacing
#   - \usepackage{float}

author:
  - name:         Natsis Athanasios
    email:        natsisphysicist@gmail.com
    affiliation:  Laboratory of Atmospheric Physics
#    footnote:     1
    correspondingauthor: true
  - name:         Alkiviadis Bais
    email:        abais@auth.gr 
    affiliation:  Laboratory of Atmospheric Physics
    correspondingauthor: false
address:
  - code:         Laboratory of Atmospheric Physics
    organization: Aristotle University of Thessaloniki
    addressline:  Campus Box 149
    city:         Thessaloniki
    # state:        State
    postcode:     54124
    country:      Greece
#  - code:         Another University
#    organization: Department
#    addressline:  A street 29
#    postcode:     2054 NX
#    city:         Manchester,
#    country:      The Netherlands
# footnote:
#   - code:         1
#     text:         "This is the first author footnote."
#   - code:         2
#     text:         "Another author footnote."
abstract: |
  Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.

  It consists of two paragraphs.
keywords: 
  - GHI
  - SDR
  - Solar Brigthening/Dimming
journal:        "An awesome journal"
date:           "`r Sys.Date()`"
classoption:    preprint, 3p, authoryear
bibliography:   references.bib
linenumbers:    false
numbersections: true

output: 
  rticles::elsevier_article:
    keep_tex:         false
    citation_package: natbib
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         yes
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
# date:           "`r format(Sys.time(), '%F')`"
---

@@

Before further analysis, we implement an appropriate aggregation scheme on the 1-minute data, to preserve the ....representabiliyt....
When we use daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
Accordingly, monthly values are computed from the daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
The daily seasonal values are indexed based on the day of year number, and the monthly by the corresponding calendar month.

<!-- Seasonal -->
For the seasons of the year trends, we also grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)

<!-- Deseasonalization -->
Finally, for each data set, we remove the natural occurring seasonal variation.
This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
Thus, for each data set we have obtained the relative departure of the seasonal mean.
These values are presented in the following section.

@@

## Aggregation

@@

Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions, the statistical significance of the trends is acceptable (Tab. \ref{tab:trendtable}).

```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The straight black line shows the long term trend for All-sky conditions."}
knitr::include_graphics("./images/LongtermTrends-2.pdf")
```

@@

```{r trendCLOUD, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for cloud sky conditions."}
knitr::include_graphics("./images/LongtermTrends-6.pdf")
```

<!-- \footnotesize -->
```{r trendtable, echo=F}
pp <- read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE)
pp <- pp[grep("Global" ,pp$var), ]
pp$slope.sd           <- NULL
pp$slope.ConfInt_0.95 <- NULL
pp$slope.ConfInt_0.99 <- NULL
pp$var                <- NULL
rownames(pp) <- NULL
names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")

panderOptions("table.alignment.default", "left")
pander(pp,
       caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
# \\label{tab:tab2}
# knitr::kable(pp, caption = "Trends of daily means by sky conditions.")
```
<!-- \normalsize -->

@@

\footnotesize
```{r trendseasontable, echo=F}
pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE)
pp <- pp[grep("Global" ,pp$var), ]
pp$slope.sd           <- NULL
pp$slope.ConfInt_0.95 <- NULL
pp$slope.ConfInt_0.99 <- NULL
pp$var                <- NULL
rownames(pp) <- NULL
names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Season", "Trend statistical sign.")

# panderOptions("table.alignment.default", "left")
panderOptions("table.alignment.default", "center")
pander(pp,
       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the season of the year.")
```
\normalsize

@@

Our result reaffirm the

>>>>

---
title:         "Trends of SDR in Thessaloniki"
# header-includes:
author:
  - name:         Natsis Athanasios
    email:        natsisphysicist@gmail.com
    affiliation:  Laboratory of Atmospheric Physics
#    footnote:     1
    correspondingauthor: true
  - name:         Alkiviadis Bais
    email:        abais@auth.gr 
    affiliation:  Laboratory of Atmospheric Physics
    correspondingauthor: false
address:
  - code:         Laboratory of Atmospheric Physics
    organization: Aristotle University of Thessaloniki
    addressline:  Campus Box 149
    city:         Thessaloniki
    # state:        State
    postcode:     54124
    country:      Greece
#  - code:         Another University
#    organization: Department
#    addressline:  A street 29
#    postcode:     2054 NX
#    city:         Manchester,
#    country:      The Netherlands
# footnote:
#   - code:         1
#     text:         "This is the first author footnote."
#   - code:         2
#     text:         "Another author footnote."
abstract: |
  Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.
  It consists of two paragraphs.
keywords: 
  - GHI
  - SDR
  - Solar Brigthening/Dimming
journal:        "An awesome journal"
date:           "`r Sys.Date()`"
classoption:    preprint, 3p, authoryear
bibliography:   references.bib
linenumbers:    false
numbersections: true
output: 
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         yes
    latex_engine:     xelatex
    toc:              yes
    toc_depth:        4
    fig_width:        7
    fig_height:       4.5
    documentclass:    article
    classoption:      a4paper,oneside
    fontsize:         12pt
    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
    link-citations:   yes
    colorlinks:       yes
    urlcolor:         blue
    biblio-style:     apalike
  rticles::elsevier_article:
    keep_tex:         false
    citation_package: natbib
# date:           "`r format(Sys.time(), '%F')`"
---

@@

<!-- aggregation -->
Before further analysis and Deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
The daily seasonal values are indexed based on the day of year number, and the monthly by the corresponding calendar month.
<!-- Seasonal -->
For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
<!-- Deseasonalization -->
Finally, for each data set, we remove the natural occurring seasonal variation.
This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
Thus, for each data set we have obtained the relative departure of the seasonal mean.
These values are presented in the following section.

@@

## SZA Aggregation

@@

Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions, the statistical significance of the trends are acceptable (Tab. \ref{tab:trendtable}), to draw some conclusions.

```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
knitr::include_graphics("./images/LongtermTrends-2.pdf")
```

@@

```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."}
knitr::include_graphics("./images/LongtermTrends-6.pdf")
```

<!-- \footnotesize -->
```{r trendtable, echo=F}
LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE))
LT <- LT[grep("Global", LT$var), ]
LT$slope.sd           <- NULL
LT$slope.ConfInt_0.95 <- NULL
LT$slope.ConfInt_0.99 <- NULL
LT$var                <- NULL
setorder(LT, -slope.stat_sig)
rownames(LT) <- NULL
names(LT) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")

panderOptions("table.alignment.default", "center")
pander(LT,
       caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
# \\label{tab:tab2}
# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")

## some vars to use
alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend %/year`], 2)
cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend %/year`], 2)
cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend %/year`], 2)
```
<!-- \normalsize -->

@@

<!-- \footnotesize -->
```{r trendseasontable, echo=F}
pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
               strip.white = TRUE,
               sep = ";",
               header = TRUE)
pp <- pp[grep("Global" ,pp$var), ]
pp$slope.sd           <- NULL
pp$slope.ConfInt_0.95 <- NULL
pp$slope.ConfInt_0.99 <- NULL
pp$var                <- NULL
rownames(pp) <- NULL
names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Season", "Trend statistical sign.")

# panderOptions("table.alignment.default", "left")
panderOptions("table.alignment.default", "center")
pander(pp,
       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year.")
```
<!-- \normalsize -->

@@

Our result for all-sky condition ($`r alltrend`\%/year$), reaffirm the previous results of @Bais2013 for the site.
The increasing of this trend shows that the phenomenon and probably the causes have been amplified.

.....
Also, the trend of $`r cloudtrend`\%/year$ for cloud-sky condition indicate the major part that clouds play on the
.....

.......
